Data-driven Reduction of Multiscale Stochastic Dynamical Systems \thanks{This work was supported by ....}

### 2014-09-01 reducible_SDE/paper/SIADS_draft.tex
added more to paper (more math)

We will consider two specific examples. 
%
In the first example, our data will form a strip where the fast and slow remain uncoupled,
\begin{equation}
\mathbf{f_1}(x_1, x_2) = \begin{bmatrix} x_1 \\ x_2 \end{bmatrix}
\end{equation}

@@

Then, \eqref{eq:general_SDE} becomes
\begin{equation} 
\begin{aligned}
d\xi_i &= b_i(\xi_1, \dots, \xi_n) dt + dW_i, \: 1 \le i \le m \\
d\xi_i &= -\frac{b_i(\xi_1, \dots, \xi_n)}{\sqrt{\epsilon}} dt + dW_i , \: m+1 \le i \le n
\end{aligned}
\end{equation}
where $b_i(\xi_1, \dots, \xi_n) = a_i (\xi_1, \dots, \xi_m, \frac{\xi_{m+1}}{\sqrt{\epsilon}}, \dots, \frac{\xi_n}{\sqrt{\epsilon}})$.
%
From \eqref{eq:general_rescale}, we can see that if $x_i = \mathcal{O}(1)$, then $\xi_i = \mathcal{O}({\sqrt{\epsilon}})$ for $ m+1 \le i \le n$, and 
\begin{equation}
\| \Xi_2 - \Xi_1 \|^2_2 = \sum_{i=1}^m \| \xi_{i,2} - \xi_{i,1} \|^2 = \sum_{i=1}^m \|x_{i,2} - x_{i,1} \|^2 + \epsilon \sum_{i=m+1}^n \|x_{i,2} - x_{i,1}\|^2
\end{equation}
%
Remarkably, this metric induces two important features. 
%
One, rescaling the data so that all variables have unit diffusion terms means that fast variables are collapsed and become ``epsilon" small, as observed in the second term in the right hand side. 
%
Two, the righthandside of \eqref{eq:mahalanobis} is invariant (to fourth order) to the function $\mathbf{f}$ and allows for the approximation of the Euclidean distance between samples of the slow variables in the ``decoupled" SDE space.

@@

Let $g = f^{-1}: \mathbb{R}^m \mapsto \mathbb{R}^n$, and let $Y_1 = f(X_1)$, and $Y_2 = f(X_2)$.
%
Again, we let $g_j^i = \frac{\partial g^i}{\partial Y^j}$.
%
Then, by Taylor expansion, we obtain
\begin{eqnarray}
X_2^i &=& X_1^i + \sum_j g_j^i (Y_1) (Y^j_2 - Y^j_1 ) 
+ \frac{1}{2} \sum_{kl}  g^i_{kl} (Y_1) (Y^k_2 - Y^k_1)(Y^l_2 - Y^l_1) \\
&&+ \frac{1}{6} \sum_{klm}  g^i_{klm} (Y_1) (Y^k_2 - Y^k_1)(Y^l_2 - Y^l_1) (Y^m_2 - Y^m_1) 
+ \mathcal{O}( \|Y_2 - Y_1\|^4 )
\end{eqnarray}

Subsitituting the Taylor expansion into the quadratic form, we obtain
\begin{eqnarray}
&&\| X_2 - X_1 \|^2 \\
&=& \frac{1}{2} (Y_2 - Y_1 )^T ((J J^T)^{-1} (Y_1) + (J J^T)^{-1}(Y_2)) (Y_2 - Y_1 ) \\
&& + \frac{1}{2} \sum_{ijkl} \left( g_j^i (Y_1) g^i_{kl} (Y_1) - g_j^i (Y_2) g^i_{kl} (Y_2) \right) (Y^j_2 - Y^j_1 )  (Y^k_2 - Y^k_1)(Y^l_2 - Y^l_1) \\
&& + \frac{1}{8} \sum_{ijklm}  \left( g^i_{jk} (Y_1) g^i_{lm} (Y_1) + g^i_{jk} (Y_2) g^i_{lm} (Y_2)  \right) (Y^j_2 - Y^j_1) (Y^k_2 - Y^k_1) (Y^l_2 - Y^l_1) (Y^m_2 - Y^m_1) \\
&& + \frac{1}{6} \sum_{ijklm}  \left( g^i_{j} (Y_1) g^i_{klm} (Y_1) + g^i_{j} (Y_2) g^i_{klm} (Y_2)  \right)(Y^j_2 - Y^j_1) (Y^k_2 - Y^k_1) (Y^l_2 - Y^l_1) (Y^m_2 - Y^m_1) \\
&& + \mathcal{O} (\|Y_1 - Y_2 \|^6 ) 
\end{eqnarray}

In general, we do not have access to $f$, $g$, or any of its derivatives.
%
However, we note that $J J^T(Y_1)$ is the local covariance of $Y_1$, which we can empirically estimate from data (this will be discussed in a subsequent section).
%
Therefore, we choose to truncate the distance approximation at the first order term.
%
We call this distance the {\em Mahalanobis distance}, 
\begin{equation}
 \| Y_2 - Y_1 \|^2_M = \frac{1}{2} (Y_2 - Y_1 )^T ((J J^T)^{-1} (Y_1) + (J J^T)^{-1}(Y_2)) (Y_2 - Y_1 ) 
\end{equation}
%
The error in the approximation is defined as
\begin{equation}
e_M(Y_1, Y_2) = \| X_2 - X_1 \|^2_2 - \| Y_2 - Y_1 \|^2_M 
\end{equation}

@@

For our general multi scale + function SDE formulation,

For our general multi scale + function SDE formulation, 
\begin{equation}
\begin{aligned}
&\mathbb{E} \left[ f_1(X_\tau) f_2(X_\tau) \right] - \mathbb{E}[f_1(X_\tau)]\mathbb{E}[f_2(X_\tau)] = \\
& \sum_{i=1}^m \left. \frac{\partial f_1}{\partial x_i} \right|_{X_\rho} \left. \frac{\partial f_2}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left( \int_\rho^\tau dW_s^i \right)^2 \right] \\
& + \frac{1}{\epsilon} \sum_{i=m+1}^n \left. \frac{\partial f_1}{\partial x_i} \right|_{X_\rho} \left. \frac{\partial f_2}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left( \int_\rho^\tau dW_s^i \right)^2 \right] \\
& + \sum_{i=1}^m \left. \frac{\partial f_1}{\partial x_i} \right|_{X_\rho} \mathbb{E} \left[ \left( \int_\rho^\tau dW_s^i  \right) \left( \int_\rho^\tau \int_\rho^{s_2} \frac{\partial^2 f_2}{\partial x_i^2} dW_{s_1}^i dW_{s_2}^i  \right) \right] \\
& + \frac{1}{\epsilon^{3/2}} \sum_{i=m+1}^n \left. \frac{\partial f_1}{\partial x_i} \right|_{X_\rho} \mathbb{E} \left[ \left( \int_\rho^\tau dW_s^i  \right) \left( \int_\rho^\tau \int_\rho^{s_2} \frac{\partial^2 f_2}{\partial x_i^2} dW_{s_1}^i dW_{s_2}^i  \right) \right] \\
& + \sum_{i=1}^m \left. \frac{\partial f_2}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left( \int_\rho^\tau dW_s^i \right)  \left( \int_\rho^\tau \int_\rho^{s_2} \frac{\partial^2 f_1}{\partial x_i^2}  dW_{s_1}^i dW_{s_2}^i \right) \right] \\
& + \frac{1}{\epsilon^{3/2}} \sum_{i=m+1}^n \left. \frac{\partial f_2}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left( \int_\rho^\tau dW_s^i \right)  \left( \int_\rho^\tau \int_\rho^{s_2} \frac{\partial^2 f_1}{\partial x_i^2}  dW_{s_1}^i dW_{s_2}^i \right) \right] \\
&+ \mathcal{O} ((\tau - \rho)^2 ) 
\end{aligned}
\end{equation}

For the case when $f$ is linear,
\begin{equation}
\begin{aligned}
& \mathbb{E} [f_1(X_\tau) f_2(X_\tau) ] - \mathbb{E}[f_1(X_\tau)]\mathbb{E}[f_2(X_\tau)]= \\
=  &  \sum_{i=1}^m  \left. \frac{\partial f_1}{\partial x_i} \right|_{X_\rho} \left. \frac{\partial f_2}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left(\int_\rho^\tau dW_s^i \right)^2  \right] \\
& + \frac{1}{\epsilon} \sum_{i=m+1}^n  \left. \frac{\partial f_1}{\partial x_i} \right|_{X_\rho} \left. \frac{\partial f_2}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left(\int_\rho^\tau dW_s^i \right)^2  \right] \\
 & + \sum_{i=1}^m \left. \frac{\partial f_1}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left(\int_\rho^\tau dW_s^i \right)  \left(\int_\rho^\tau \int_{\rho}^{s_2} \left(  \sum_{k=1}^m \frac{\partial a_k}{\partial x_i} \frac{\partial f_2}{\partial x_k}  + \frac{1}{\epsilon} \sum_{k=m+1}^m \frac{\partial a_k }{\partial x_i} \frac{\partial f_2}{\partial x_k} \right)  dW_{s_1}^i ds_2 \right) \right] \\ 
 & + \frac{1}{\epsilon} \sum_{i=m+1}^n \left. \frac{\partial f_1}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left(\int_\rho^\tau dW_s^i \right)  \left(\int_\rho^\tau \int_{\rho}^{s_2} \left(  \sum_{k=1}^m \frac{\partial a_k}{\partial x_i} \frac{\partial f_2}{\partial x_k}  + \frac{1}{\epsilon} \sum_{k=m+1}^m \frac{\partial a_k }{\partial x_i} \frac{\partial f_2}{\partial x_k} \right)  dW_{s_1}^i ds_2 \right) \right] \\ 
 & + \sum_{i=1}^m  \left. \frac{\partial f_2}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left(\int_\rho^\tau dW_s^i \right)  \left(\int_\rho^\tau \int_{\rho}^{s_2} \left(  \sum_{k=1}^m \frac{\partial a_k}{\partial x_i} \frac{\partial f_1}{\partial x_k}  + \frac{1}{\epsilon} \sum_{k=m+1}^m \frac{\partial a_k }{\partial x_i} \frac{\partial f_1}{\partial x_k} \right) dW_{s_1}^i ds_2 \right) \right] \\
& + \frac{1}{\epsilon} \sum_{i=m+1}^n  \left. \frac{\partial f_2}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left(\int_\rho^\tau dW_s^i \right)  \left(\int_\rho^\tau \int_{\rho}^{s_2} \left(  \sum_{k=1}^m \frac{\partial a_k}{\partial x_i} \frac{\partial f_1}{\partial x_k}  + \frac{1}{\epsilon} \sum_{k=m+1}^m \frac{\partial a_k }{\partial x_i} \frac{\partial f_1}{\partial x_k} \right) dW_{s_1}^i ds_2 \right) \right] \\
& + \mathcal{O} ((\tau - \rho)^{5/2})
 \end{aligned}
\end{equation}

Therefore, we have
\begin{equation}
\begin{aligned}
e_c = &  \sum_{i=1}^m \left. \frac{\partial f_1}{\partial x_i} \right|_{X_\rho} \mathbb{E} \left[ \left( \int_\rho^\tau dW_s^i  \right) \left( \int_\rho^\tau \int_\rho^{s_2} \frac{\partial^2 f_2}{\partial x_i^2} dW_{s_1}^i dW_{s_2}^i  \right) \right] \\
& + \frac{1}{\epsilon^{3/2}} \sum_{i=m+1}^n \left. \frac{\partial f_1}{\partial x_i} \right|_{X_\rho} \mathbb{E} \left[ \left( \int_\rho^\tau dW_s^i  \right) \left( \int_\rho^\tau \int_\rho^{s_2} \frac{\partial^2 f_2}{\partial x_i^2} dW_{s_1}^i dW_{s_2}^i  \right) \right] \\
& + \sum_{i=1}^m \left. \frac{\partial f_2}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left( \int_\rho^\tau dW_s^i \right)  \left( \int_\rho^\tau \int_\rho^{s_2} \frac{\partial^2 f_1}{\partial x_i^2}  dW_{s_1}^i dW_{s_2}^i \right) \right] \\
& + \frac{1}{\epsilon^{3/2}} \sum_{i=m+1}^n \left. \frac{\partial f_2}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left( \int_\rho^\tau dW_s^i \right)  \left( \int_\rho^\tau \int_\rho^{s_2} \frac{\partial^2 f_1}{\partial x_i^2}  dW_{s_1}^i dW_{s_2}^i \right) \right] \\
&+ \mathcal{O} ((\tau - \rho)^2 ) 
\end{aligned}
\end{equation}
%
when $\frac{\partial f}{\partial x_i} \ne 0$, and 
%
\begin{equation}
\begin{aligned}
e_c =  &  \sum_{i=1}^m \left. \frac{\partial f_1}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left(\int_\rho^\tau dW_s^i \right)  \left(\int_\rho^\tau \int_{\rho}^{s_2} \left(  \sum_{k=1}^m \frac{\partial a_k}{\partial x_i} \frac{\partial f_2}{\partial x_k}  + \frac{1}{\epsilon} \sum_{k=m+1}^m \frac{\partial a_k }{\partial x_i} \frac{\partial f_2}{\partial x_k} \right)  dW_{s_1}^i ds_2 \right) \right] \\ 
 & + \frac{1}{\epsilon} \sum_{i=m+1}^n \left. \frac{\partial f_1}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left(\int_\rho^\tau dW_s^i \right)  \left(\int_\rho^\tau \int_{\rho}^{s_2} \left(  \sum_{k=1}^m \frac{\partial a_k}{\partial x_i} \frac{\partial f_2}{\partial x_k}  + \frac{1}{\epsilon} \sum_{k=m+1}^m \frac{\partial a_k }{\partial x_i} \frac{\partial f_2}{\partial x_k} \right)  dW_{s_1}^i ds_2 \right) \right] \\ 
 & + \sum_{i=1}^m  \left. \frac{\partial f_2}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left(\int_\rho^\tau dW_s^i \right)  \left(\int_\rho^\tau \int_{\rho}^{s_2} \left(  \sum_{k=1}^m \frac{\partial a_k}{\partial x_i} \frac{\partial f_1}{\partial x_k}  + \frac{1}{\epsilon} \sum_{k=m+1}^m \frac{\partial a_k }{\partial x_i} \frac{\partial f_1}{\partial x_k} \right) dW_{s_1}^i ds_2 \right) \right] \\
& + \frac{1}{\epsilon} \sum_{i=m+1}^n  \left. \frac{\partial f_2}{\partial x_i} \right|_{X_\rho}  \mathbb{E} \left[ \left(\int_\rho^\tau dW_s^i \right)  \left(\int_\rho^\tau \int_{\rho}^{s_2} \left(  \sum_{k=1}^m \frac{\partial a_k}{\partial x_i} \frac{\partial f_1}{\partial x_k}  + \frac{1}{\epsilon} \sum_{k=m+1}^m \frac{\partial a_k }{\partial x_i} \frac{\partial f_1}{\partial x_k} \right) dW_{s_1}^i ds_2 \right) \right] \\
& + \mathcal{O} ((\tau - \rho)^{5/2})
\end{aligned}
\end{equation}

@@

>>>>

We will consider two specific examples. 
%
In the first example, our transformation $\mathbf{f}$ will be the identity function and fast and slow remain uncoupled,
\begin{equation}
\mathbf{f_1}(x_1, x_2) = \begin{bmatrix} x_1 \\ x_2 \end{bmatrix}
\end{equation}

@@

Then, \eqref{eq:general_SDE} becomes
\begin{equation} 
\begin{aligned}
d\xi_i &= b_i(\xi_1, \dots, \xi_n) dt + dW_i, \: 1 \le i \le m \\
d\xi_i &= -\frac{b_i(\xi_1, \dots, \xi_n)}{\sqrt{\epsilon}} dt + dW_i , \: m+1 \le i \le n
\end{aligned}
\end{equation}
where $b_i(\xi_1, \dots, \xi_n) = a_i (\xi_1, \dots, \xi_m, \frac{\xi_{m+1}}{\sqrt{\epsilon}}, \dots, \frac{\xi_n}{\sqrt{\epsilon}})$.
%
From \eqref{eq:general_rescale}, we can see that if $x_i = \mathcal{O}(1)$, then $\xi_i = \mathcal{O}({\sqrt{\epsilon}})$ for $ m+1 \le i \le n$, and 
\begin{equation}
\| \Xi_2 - \Xi_1 \|^2_2 = \sum_{i=1}^m \| \xi_{2,i} - \xi_{1,i} \|^2 = \sum_{i=1}^m \|x_{2,i} - x_{1,i} \|^2 + \epsilon \sum_{i=m+1}^n \|x_{2,i} - x_{1,i}\|^2
\end{equation}
%
Remarkably, this metric induces two important features. 
%
One, rescaling the data so that all variables have unit diffusion terms means that fast variables are collapsed and become ``epsilon" small, as observed in the second term in the right hand side. 
%
Two, the righthandside of \eqref{eq:mahalanobis} is invariant (to fourth order) to the function $\mathbf{f}$ and allows for the approximation of the Euclidean distance between samples of the slow variables in the ``decoupled" SDE space.

@@

Let $\mathbf{g} = \mathbf{f}^{-1}: \mathbb{R}^d \mapsto \mathbb{R}^n$, and let $Y_1 = f(X_1)$, and $Y_2 = f(X_2)$.
%
Again, we let $g_j^i = \frac{\partial g^i}{\partial Y^j}$.
%
Then, by Taylor expansion, we obtain
\begin{equation}
\begin{aligned}
x_{2,i} &=& 
x_{1,i} 
+ \sum_{j=1}^d \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_1} (y_{2,j} - y_{1,j} ) 
+ \frac{1}{2} \sum_{j=1}^d \sum_{k=1}^d \left. \frac{\partial^2 g_i}{\partial y_j \partial y_k} \right|_{Y_1} (y_{2,j} - y_{1,j}) (y_{2,k} - y_{1,k}) \\
&&+ \frac{1}{6} \sum_{j=1}^d \sum_{k=1}^d \sum_{l=1}^d \left. \frac{\partial^3 g_i}{\partial y_j \partial y_k \partial y_l} \right|_{Y_1} (y_{2,j} - y_{1,j}) (y_{2,k} - y_{1,k}) (y_{2,l} - y_{1,l})
+ \mathcal{O}( \|Y_2 - Y_1\|^4 )
\end{aligned}
\end{equation}

Substituting the Taylor expansion into the quadratic form, we obtain
\begin{equation}
\begin{aligned}
\| X_2 - X_1 \|^2_2 
=& \frac{1}{2} (Y_2 - Y_1 )^T \left(\left.(J J^T)^{-1}\right|_{Y_1} + \left.(J J^T)^{-1}\right|_{Y_2}\right) (Y_2 - Y_1 ) \\
& + \frac{1}{2} \sum_{i=1}^n \sum_{jkl=1}^{d} \left( \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_1} \left. \frac{\partial^2 g_i}{\partial y_k \partial y_l} \right|_{Y_1} - \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_2} \left. \frac{\partial^2 g_i}{\partial y_k \partial y_l} \right|_{Y_2} \right) (y_{2,j} - y_{1,j})  (y_{2,k} - y_{1,k})(y_{2,l} - y_{1,l}) \\
& + \frac{1}{8} \sum_{i=1}^n \sum_{jklm=1}^d  \left( \left. \frac{\partial^2 g_i}{\partial y_j \partial y_k} \right|_{Y_1} \left. \frac{\partial^2 g_i}{\partial y_l \partial y_m} \right|_{Y_1} + \left. \frac{\partial^2 g_i}{\partial y_j \partial y_k} \right|_{Y_2} \left. \frac{\partial^2 g_i}{\partial y_l \partial y_m} \right|_{Y_2} \right) (y_{2,j} - y_{1,j})  (y_{2,k} - y_{1,k})(y_{2,l} - y_{1,l})(y_{2,m} - y_{1,m}) \\
& + \frac{1}{6}  \sum_{i=1}^n \sum_{jklm=1}^d  \left( \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_1} \left. \frac{\partial^3 g_i}{ \partial y_k \partial y_l \partial y_m} \right|_{Y_1} + \left. \frac{\partial g_i}{\partial y_j} \right|_{Y_2} \left. \frac{\partial^3 g_i}{ \partial y_k \partial y_l \partial y_m} \right|_{Y_2} \right) (y_{2,j} - y_{1,j})  (y_{2,k} - y_{1,k})(y_{2,l} - y_{1,l})(y_{2,m} - y_{1,m}) \\
& + \mathcal{O} (\|Y_1 - Y_2 \|^6 ) 
\end{aligned}
\end{equation}

In general, we do not have access to $f$, $g$, or any of its derivatives.
%
However, we note that $J J^T(Y_1)$ is the local covariance of $Y_1$, which we can empirically estimate from data (this will be discussed in a subsequent section).
%
Therefore, we choose to truncate the distance approximation at the first order term.
%
We call this distance the {\em Mahalanobis distance}, 
\begin{equation}
 \| Y_2 - Y_1 \|^2_M = \frac{1}{2} (Y_2 - Y_1 )^T \left(\left.(J J^T)^{-1}\right|_{Y_1} + \left.(J J^T)^{-1}\right|_{Y_2}\right) (Y_2 - Y_1 )
\end{equation}
%
The error in the approximation is defined as
\begin{equation}
e_M(Y_1, Y_2) = \| X_2 - X_1 \|^2_2 - \| Y_2 - Y_1 \|^2_M 
\end{equation}

@@

To compute the Mahalanobis distance, we require $J J^T$. 
%
From \cite{...}, we know that $JJ^T = C$, where $C$ is the local covariance of the observed stochastic process. 
%
Therefore, we would like to estimate the local covariance at a point $X_t$.

We know that, for a stochastic process following \eqref{eq:general_SDE} and observed through a function $\mathbf{f}$, the covariance $C$ is given by 
\begin{equation}
C_{jk}(X_t) = 
\sum_{i=1}^m \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t} \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t} 
+ \frac{1}{\epsilon} \sum_{i=m+1}^n \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t} \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t} 
\end{equation}

We will use local simulation bursts to estimate the local covariance.
%
Let $X_\tau$ denote the samples at time $\tau > t$. 
%
For the general SDE formulation outlined in \eqref{eq:general_SDE}, we write the estimated covariance $\hat{C}$ as
\begin{equation}
\begin{aligned}
\hat{C}_{jk} (X_t, \tau - t) = & 
\frac{1}{\tau - t} \left(  \mathbb{E} \left[ f_j(X_\tau) f_k(X_\tau) \right] - \mathbb{E}[f_j(X_\tau)]\mathbb{E}[f_k(X_\tau)]   \right)\\
= & \sum_{i=1}^m \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t} \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t}  \\
& + \frac{1}{\epsilon} \sum_{i=m+1}^n \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t} \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t} \\
& + \frac{1}{\tau - t} \sum_{i=1}^m \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t} \mathbb{E} \left[ \left( \int_t^\tau dW_s^i  \right) \left( \int_t^\tau \int_t^{s_2} \frac{\partial^2 f_k}{\partial x_i^2} dW_{s_1}^i dW_{s_2}^i  \right) \right] \\
& + \frac{1}{\epsilon^{3/2} (\tau - t)} \sum_{i=m+1}^n \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t} \mathbb{E} \left[ \left( \int_t^\tau dW_s^i  \right) \left( \int_t^\tau \int_t^{s_2} \frac{\partial^2 f_k}{\partial x_i^2} dW_{s_1}^i dW_{s_2}^i  \right) \right] \\
& + \frac{1}{\tau - t} \sum_{i=1}^m \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left( \int_t^\tau dW_s^i \right)  \left( \int_t^\tau \int_t^{s_2} \frac{\partial^2 f_j}{\partial x_i^2}  dW_{s_1}^i dW_{s_2}^i \right) \right] \\
& + \frac{1}{\epsilon^{3/2} (\tau - t)} \sum_{i=m+1}^n \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left( \int_t^\tau dW_s^i \right)  \left( \int_t^\tau \int_t^{s_2} \frac{\partial^2 f_j}{\partial x_i^2}  dW_{s_1}^i dW_{s_2}^i \right) \right] \\
&+ \mathcal{O} ((\tau - t) ) 
\end{aligned}
\end{equation}
%
However, for the specific case when $\mathbf{f}$ is linear, the $\mathcal{O}((\tau - t)^{3/2})$ terms go to 0, and we must consider higher-order terms. 
%
In this case, 
\begin{equation}
\begin{aligned}
\hat{C}_{jk} (X_t, \tau - t) = & 
\frac{1}{\tau - t} \left(  \mathbb{E} \left[ f_j(X_\tau) f_k(X_\tau) \right] - \mathbb{E}[f_j(X_\tau)]\mathbb{E}[f_k(X_\tau)]   \right)\\
= & \sum_{i=1}^m \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t} \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t}  \\
& + \frac{1}{\epsilon} \sum_{i=m+1}^n \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t} \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t} \\
 & + \frac{1}{\tau - t} \sum_{i=1}^m \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^i \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \sum_{l=1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_k}{\partial x_l}  + \frac{1}{\epsilon} \sum_{l=m+1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_k}{\partial x_l} \right)  dW_{s_1}^i ds_2 \right) \right] \\ 
 & + \frac{1}{\epsilon (\tau - t)} \sum_{i=m+1}^n \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^i \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \sum_{l=1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_k}{\partial x_l}  + \frac{1}{\epsilon} \sum_{l=m+1}^m \frac{\partial a_l }{\partial x_i} \frac{\partial f_k}{\partial x_l} \right)  dW_{s_1}^i ds_2 \right) \right] \\ 
 & + \frac{1}{\tau - t} \sum_{i=1}^m  \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^i \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \sum_{l=1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_j}{\partial x_l}  + \frac{1}{\epsilon} \sum_{l=m+1}^m \frac{\partial a_l }{\partial x_i} \frac{\partial f_j}{\partial x_l} \right) dW_{s_1}^i ds_2 \right) \right] \\
& + \frac{1}{\epsilon (\tau - t)} \sum_{i=m+1}^n  \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^i \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \sum_{l=1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_j}{\partial x_l}  + \frac{1}{\epsilon} \sum_{l=m+1}^m \frac{\partial a_l }{\partial x_i} \frac{\partial f_j}{\partial x_l} \right) dW_{s_1}^i ds_2 \right) \right] \\
& + \mathcal{O} ((\tau - t)^{3/2})
 \end{aligned}
\end{equation}
%
For both cases,  we define the error in the estimation as
%
\begin{equation}
\begin{aligned}
e_C (X_t, \tau - t) = \| \hat{C} (X_t, \tau - t)  - C(X_t, \tau - t) \|
\end{aligned}
\end{equation}

@@

For illustrative purposes, we will consider the following two-dimensional SDE as a specific example of \eqref{eq:general_SDE}. 
\begin{equation}
\begin{aligned}
dx_1 &= adt + dW_1\\
dx_2 &= -\frac{x_2}{\epsilon} dt + \frac{1}{\sqrt{\epsilon}} dW_2
\end{aligned}
\end{equation}
%
where $a$ is a constant of order 1 (we will take $a=3$). 
%
So $x_1$ is the slow variable, and $x_2$ is a fast noise whose equilibrium measure is bounded and $\mathcal{O}(1)$.
%
We would like to recover $x_1$ using data-driven techniques.

\subsection{Linear functions}

In the first example, our transformation $\mathbf{f}$ will be the identity function and fast and slow remain uncoupled,
\begin{equation}
\mathbf{f}(x_1, x_2) = \begin{bmatrix} x_1 \\ x_2 \end{bmatrix}
\end{equation}
%
and
%
\begin{equation}
\mathbf{g}(y_1, y_2) = \mathbf{f}^{-1} (y_1, y_2) = \begin{bmatrix} y_1 \\ y_2 \end{bmatrix}
\end{equation}

In this case, $e_M = 0$, as the second- and higher-order derivatives of $\mathbf{g}$ are all 0.

\begin{equation}
\begin{aligned}
e_{C, jk} (X_t, \tau - t) = & 
\frac{1}{\tau - t} \sum_{i=1}^m \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^i \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \sum_{l=1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_k}{\partial x_l}  + \frac{1}{\epsilon} \sum_{l=m+1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_k}{\partial x_l} \right)  dW_{s_1}^i ds_2 \right) \right] \\ 
 & + \frac{1}{\epsilon (\tau - t)} \sum_{i=m+1}^n \left. \frac{\partial f_j}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^i \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \sum_{l=1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_k}{\partial x_l}  + \frac{1}{\epsilon} \sum_{l=m+1}^m \frac{\partial a_l }{\partial x_i} \frac{\partial f_k}{\partial x_l} \right)  dW_{s_1}^i ds_2 \right) \right] \\ 
 & + \frac{1}{\tau - t} \sum_{i=1}^m  \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^i \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \sum_{l=1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_j}{\partial x_l}  + \frac{1}{\epsilon} \sum_{l=m+1}^m \frac{\partial a_l }{\partial x_i} \frac{\partial f_j}{\partial x_l} \right) dW_{s_1}^i ds_2 \right) \right] \\
& + \frac{1}{\epsilon (\tau - t)} \sum_{i=m+1}^n  \left. \frac{\partial f_k}{\partial x_i} \right|_{X_t}  \mathbb{E} \left[ \left(\int_t^\tau dW_s^i \right)  \left(\int_t^\tau \int_{t}^{s_2} \left(  \sum_{l=1}^m \frac{\partial a_l}{\partial x_i} \frac{\partial f_j}{\partial x_l}  + \frac{1}{\epsilon} \sum_{l=m+1}^m \frac{\partial a_l }{\partial x_i} \frac{\partial f_j}{\partial x_l} \right) dW_{s_1}^i ds_2 \right) \right] \\
& + \mathcal{O} ((\tau - t)^{3/2})
 \end{aligned}
\end{equation}

\subsection{Nonlinear functions}

In the second example, our data will be warped into half-moon shapes
\begin{equation}
\mathbf{f_2}(x_1, x_2) = 
\begin{bmatrix} 
x_1 + x_2^2 \\
x_2
\end{bmatrix}
\end{equation}
